### 2021-02-09 final_report/include/CAMT.tex
reviewed MT eval chapter

Automating CA can be seen, then, as a first step for automatically generating such lexicon. 
An important point is that our approach, as opposed to word alignment techniques, identifies concepts at different levels of abstraction.
This allows us to construct lexical entries not only for the grammatical categories that typically correspond to single words (such as nouns, adjectives and adverbs), but also for phrases (with the term ``phrase'' now intended grammatically, as opposed to its meaning in SMT).
Doing so enables GF translation, designed to preserve grammatically, to also render idiomatic expressions found in the data used for generating the lexicon correctly. \smallskip

@@

\subsection{UD-to-GF conversion}
The output of both CE and CP is, unless when linearized for evaluation purposes, a set of aligned UD trees in CoNNL-U format. In order to be used for GF-based translation, such trees need to be first converted into GF ASTs, which are then used to generate a set of grammar rules constiuting the above mentioned multilingual GF lexicon. \smallskip

@@

which specifies that a \texttt{PredVP} (predicative Verb Phrase), defined in the RGL as a function of type

@@

A \textit{morphological dictionary}, on the other hand, contains correspondences between a large number of lemmas and various word forms. In this case, the morphological dictionaries themselves are implemented in GF. \smallskip

@@

Of course, some limited variation is still possible. For instance, the determiner can change or disappear, number can vary, the complement of the copula can become a noun and an adjectival modifiers and prepositional phrases can be added to each noun phrase. For instance:

\begin{example} \ 
    \begin{itemize}
        \item ``the sentence is simple''
        \item ``a sentence is simple''
        \item ``sentences are simple''
        \item ``these sentences are simple''
        \item ``this sentence is an example''
        \item ``this short sentence is simple''
        \item ``this sentence of the text is simple''
    \end{itemize}
\end{example}

@@

\begin{example} \
    \begin{itemize}
        \item ``this sentence isn't simple''
        \item ``is this sentence simple?''
        \item ``this sentence was simple''
        \item ``this sentence will be simple''
        \item ``this sentence is simpler than that sentence''
    \end{itemize}
\end{example}

@@

\begin{itemize}
    \item while GF can parse the SL string in a variety of ways and, as a consequence, return multiple TL linearizations, our program only outputs the first candidate translation, preliminarily discarding trees containing terminal nodes that have no linearization in the TL\footnote{a linearization can be missing, for instance, if a concrete syntax is postprocessed by simply getting rid of an ``illegal'' verb causing compilation problems (cf. Section \ref{grgn})} 
    \item to facilitate the experiments, our standalone program can translate several newline-separated sentences at once.
\end{itemize}

@@

\begin{figure}[h]
    \centering
    \footnotesize
    \begin{subfigure}{.5\textwidth}
      \centering
      \setlength{\unitlength}{0.2mm}
      \begin{picture}(315.0,110.0)
        \put(0.0,0.0){parse}
        \put(55.0,0.0){this}
        \put(101.0,0.0){sentence}
        \put(183.0,0.0){with}
        \put(229.0,0.0){care}
        \put(0.0,15.0){{\tiny VERB}}
        \put(55.0,15.0){{\tiny DET}}
        \put(101.0,15.0){{\tiny NOUN}}
        \put(183.0,15.0){{\tiny ADP}}
        \put(229.0,15.0){{\tiny NOUN}}
        \put(88.0,30.0){\oval(39.47826086956522,33.333333333333336)[t]}
        \put(68.26086956521739,35.0){\vector(0,-1){5.0}}
        \put(73.0,49.66666666666667){{\tiny det}}
        \put(70.5,30.0){\oval(98.02970297029702,66.66666666666667)[t]}
        \put(119.51485148514851,35.0){\vector(0,-1){5.0}}
        \put(55.5,66.33333333333334){{\tiny obj}}
        \put(216.0,30.0){\oval(39.47826086956522,33.333333333333336)[t]}
        \put(196.26086956521738,35.0){\vector(0,-1){5.0}}
        \put(201.0,49.66666666666667){{\tiny case}}
        \put(134.5,30.0){\oval(227.68995633187774,100.0)[t]}
        \put(248.34497816593887,35.0){\vector(0,-1){5.0}}
        \put(119.5,83.0){{\tiny obl}}
        \put(15.0,110.0){\vector(0,-1){80.0}}
        \put(20.0,100.0){{\tiny root}}
      \end{picture}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \setlength{\unitlength}{0.25mm}
\begin{picture}(324.0,90.0)
  \put(0.0,0.0){analizza}
  \put(82.0,0.0){questa}
  \put(146.0,0.0){frase}
  \put(201.0,0.0){con}
  \put(238.0,0.0){cura}
  \put(0.0,15.0){{\tiny VERB}}
  \put(82.0,15.0){{\tiny DET}}
  \put(146.0,15.0){{\tiny NOUN}}
  \put(201.0,15.0){{\tiny ADP}}
  \put(238.0,15.0){{\tiny NOUN}}
  \put(124.0,30.0){\oval(59.3125,33.333333333333336)[t]}
  \put(94.34375,35.0){\vector(0,-1){5.0}}
  \put(109.0,49.66666666666667){{\tiny det}}
  \put(93.0,30.0){\oval(143.94520547945206,66.66666666666667)[t]}
  \put(164.97260273972603,35.0){\vector(0,-1){5.0}}
  \put(78.0,66.33333333333334){{\tiny obj}}
  \put(229.5,30.0){\oval(28.89189189189189,33.333333333333336)[t]}
  \put(215.05405405405406,35.0){\vector(0,-1){5.0}}
  \put(214.5,49.66666666666667){{\tiny case}}
  \put(212.0,30.0){\oval(88.73913043478261,66.66666666666667)[t]}
  \put(256.3695652173913,35.0){\vector(0,-1){5.0}}
  \put(197.0,66.33333333333334){{\tiny nmod}}
  \put(15.0,90.0){\vector(0,-1){60.0}}
  \put(20.0,80.0){{\tiny root}}
\end{picture}
    \end{subfigure}
    \caption[A case where alignment ``by exclusion'' is beneficial]{A case where alignment ``by exclusion'' is beneficial. In this case, while the two sentences should appear to be structured identically to a human reader, the Italian parse tree contains a (reasonable) attachment error. As a consequence, vanilla CE would not be able to align the prepositional phase ``\textit{with care}'', attached to the main verb ``\textit{parse}'', to its Italian exact counterpart ``\textit{con cura}'', incorrectly linked to the noun ``\textit{frase}'' as an \texttt{nmod}. A second CE pass trying to put unaligned subtrees in relation with each other, however, is able to detect the POS-equivalence between the two phrases, thus producing three new correct alignments: $\langle$\textit{with care, con cura}$\rangle$, $\langle$\textit{care, cura}$\rangle$ and $\langle$\textit{with, con}$\rangle$.}
    \label{rest}
\end{figure}

@@

On the other hand, the concepts represented in the GF grammar should intuitively, in most cases, be ``as small as possible''. If we reconsider the example in Figure \ref{cake} (sentences \textit{``finding useful correspondences is not exactly a piece of cake''} and \textit{``trovare corrispondenze utili non è proprio scontato''}), for instance, while the multi-word correspondence $\langle$\textit{a piece of cake, scontato}$\rangle$ is indeed useful there is usually no point in keeping correspondences such as $\langle$\textit{useful correspondences, corrispondenze utili}$\rangle$, even if they are correct, when it it also possible to extract the alignments $\langle$\textit{useful, utili}$\rangle$ and $\langle$\textit{correspondences, corrispondenze}$\rangle$: one-word and, in general, shorter alignments are more easily reusable in different contexts, especially since there are language-specific RGL rules capable of handling word order properly. 
As a consequence, for the purposes of grammar generation, the CA modules should in this case only return the following set of alignments: \smallskip

\begin{enumerate}
    \item $\langle$\textit{finding, trovare}$\rangle$
    \item $\langle$\textit{useful, utili}$\rangle$
    \item $\langle$\textit{correspondences, corrispondenze}$\rangle$
    \item $\langle$\textit{is, è}$\rangle$
    \item $\langle$\textit{not, non}$\rangle$
    \item $\langle$\textit{exactly, proprio}$\rangle$
    \item $\langle$\textit{a piece of cake, scontato}$\rangle$.
\end{enumerate} \smallskip

@@

In some situations, however, larger alignments (such as $\langle$\textit{useful correspondences, corrispondenze utili}$\rangle$) are considered to be relevant since they help understand which words can be safely used in conjunction with each other. For instance, while the Italian adjective ``\textit{brutto}'' (most often translated as ``\textit{ugly}'') is often used in conjunction with the word ``\textit{notizia}'' (``\textit{news}'') in expressions such as ``\textit{brutte notizie}'', in English the adjective ``\textit{bad}'' (as in ``\textit{bad news}'') is definitely more appropriate. Discarding a potential $\langle\textit{bad news}, \textit{brutte notizie}\rangle$ alignment is then not always the right choice, since in a context in which there are many alternative translation equivalents for the word ``\textit{brutto}'' it provides useful information about what adjective to use to refer to the noun ``\textit{news}''. \smallskip

@@

Thus, the algorithm is safe - in the sense that it does not get rid of any potentially meaningful alignment - in that it gets rid of superficially perfect alignments whose subtrees are also members of smaller alignments. This means that when sentences vary widely syntax-wise, as it happens in cases where translation is not literal or in the presence of parse errors, the resulting extracted concepts can still contain a large number of nodes.

@@

In a preliminary experiment conducted in this way, for instance, the Italian translation ``\textit{il docente discute la Didattica}'' of the English sentence ``\textit{the teacher discusses the didactics}'', was given the minimum score despite being completely correct and even idiomatic\footnote{typically, the adjective ``didattica'' is not use in its plural form ``didattiche'' when nominalized. The capital D is often used, for this particular term, in official university documents.}. 
Looking at the two reference translations used at the time, it appears clear that such a low score is due to the capitalization of the word ``Didattica'' and the arbitrary (but equally correct) lexical choices made by the human translators, who rendered the sentence as ``il professore discute i metodi didattici'' and ``l' insegnante discute la didattica''\footnote{The terms ``docente'', ``professore'' and ``insegnante'' are almost perfect synonyms, even though ``docente'' is mostly used in formal documents to refer to university lecturers, ``professore'' is more common in secondary school and colloquial speech, and ``insegnante'' is the most generic. As such, it can also refer to elementary school teachers and in some cases even sport coaches. The expression ``metodi didattici'' is a perfectly sensible replacement for ``didattica''.}. 
Coincidentally (but not unlikely for a language like Italian), this is a case in which these lexical differences affect the choice of the determiners used in the various translations (cf. ``\underline{l'}insegnante'' vs. ``\underline{il} professore'' and ``\underline{la} didattica'' vs. ``\underline{i} metodi didattici''), producing a BLEU-1 to 4 score of 0. \smallskip

@@

These synthetic figures are useful to give an idea of the general quality of the translations: overall, although with relatively low scores, English-to-Swedish translation works significantly better than English-to-Italian. 
Looking back at the results obtained for CE (cf. Section \ref{eval2}), this is not excessively surprising, since both precision and recall for English-Swedish are consistently higher than they are for English-Italian. 
When it comes to the results obtained for CE on the course plans corpora in particular (cf. Table \ref{tcereal}), however, the difference in precision between the two language pairs is negligible.
This makes the reason for the substantial difference observed in the BLEU scores obtained for the two different corpora unclear, at least at a first glance.\smallskip

@@

On the other hand, it is easy for shorter sentences to be assigned the minimum BLEU-1 to 4 score even when they only contain a single grammatical or semantic error. 
This is the case, for instance, of the sentence ``\textit{the test is oral}'', whose last word, ``\textit{oral}'' is translated as ``\textit{dura}'' (``\textit{hard}'') instead of ``\textit{orale}'' due to an alignment error. 
Nonetheless, it is worth noticing that this is one of the many cases in which, even though the fact that only the first translation candidate is taken into account hides it, the correct alignment $\langle$``\textit{oral}'',``\textit{orale}''$\rangle$ is also found by the CE module. 
Moreover, the such valid correspondence has a higher number of occurrences and matches stronger alignment criteria than the wrong one $\langle$``\textit{oral}'',``\textit{duro}''$\rangle$. 
This suggest that, in contexts where higher precision is necessary, there are obvious improvements that can be done either in the choice of the alignment criteria or at a later stage, when relevant alignments are selected. \smallskip

@@

In Italian, grammar errors often involve contractions such as ``\textit{del}'' (``\textit{di}'' + ``\textit{il}'', in English ``\textit{of the}''), some of which are systematically rendered as two separate words.
Another common case is that of wrong (or at least very confusing) adjective collocation, such as in the translation ``\textit{il libro presenta una tecnica con miglioramenti vari utile}'' (``\textit{the textbook presents a useful technique with various improvements}''), where the adjective ``\textit{utile}'' (``\textit{useful}''), referred to ``\textit{technique}'' (``\textit{tecnica}'') is placed far from such noun, making the sentence hard to interpret\footnote{the manually postprocessed translation is ``\textit{il libro presenta una utile tecnica con miglioramenti vari}''.}.
Grammatical errors in Swedish are less common and, apart from long adjectives never being turned to the comparative degree periphrastically (e.g. ``\textit{relevant\underline{are}}'' instead of ``\textit{mer relevant}''), less systematic. 
Only in one case, for instance, (``\textit{programbibliotek\underline{en}}''), gender is incorrect. \smallskip

Some alignment errors are also interesting to analyze.
In English-Swedish, while several compounds, such as $\langle$\textit{computer science, datavetenskap}$\rangle$, are aligned correctly, there are cases in which a single-root noun is translated as a compound (e.g. $\langle$ \textit{theory, automatateori} $\rangle$). This is not necessarily an actual alignment error, as it might rather be that the English version of the text was being less specific than its Swedish counterpart, thus producing an alignment that, during evaluation, would have been marked as correct but not generally reusable (\texttt{=}). \smallskip

``Reasonable'' alignment errors appear in the DMI corpus sentences too. 
The English for ``\textit{attendance of lessons}'', for instance, becomes simply ``\textit{frequenza}'' in the Italian translation. 
Such word is in fact often used alone to replace longer expressions such as ``\textit{frequenza delle lezioni}'', just like in English `\textit{of lessons}'' can be omitted when what must be attended is evident from the context.
Another interesting example is the alignment $\langle$ \textit{class, classe} $\rangle$, which causes the sentence \textit{``I will attend the class''} to be (incorrectly) translated as \textit{``io seguirò la classe''} instead of \textit{``io seguirò la lezione''} even though the correspondence is in fact valid in most of the numerous contexts in which \textit{``class''} is not to be intended as a synonym of \textit{``lesson''}.

>>>>

Automating CA can be seen, then, as a first step for automatically generating such lexicon. 
An important point is that our approach, as opposed to word alignment techniques, identifies concepts at different levels of abstraction.
This allows us to construct lexical entries not only for the grammatical categories that typically correspond to single words (such as nouns, adjectives and adverbs), but also for phrases (with the term ``phrase'' now intended grammatically, as opposed to its meaning in SMT).
Doing so enables GF translation, designed to preserve grammaticality, to also render idiomatic expressions found in the data used for generating the lexicon correctly. \smallskip

@@

\subsection{UD-to-GF conversion}
The output of both CE and CP is, excepts when linearized for evaluation purposes, a set of aligned UD trees in CoNNL-U format. 
In order to be used for GF-based translation, such trees need to be first converted into GF ASTs, which are then used to generate a set of grammar rules constiuting the above mentioned multilingual GF lexicon. \smallskip

@@

which specifies that a \texttt{PredVP} (predicative Verb Phrase), defined in the RGL as of type

@@

A \textit{morphological dictionary}, on the other hand, contains correspondences between a large number of lemmas and various word forms. 
In this case, the morphological dictionaries themselves are implemented in GF, as part of the RGL. \smallskip

@@

Of course, some limited variation is still possible. 
For instance, the determiner can change or disappear, number can vary, the complement of the copula can become a noun and an adjectival modifiers and prepositional phrases can be added to each noun phrase. 
Here are some examples:

\begin{example} \ \smallskip
    \begin{itemize}
        \item ``the sentence is simple''
        \item ``a sentence is simple''
        \item ``sentences are simple''
        \item ``these sentences are simple''
        \item ``this sentence is an example''
        \item ``this short sentence is simple''
        \item ``this sentence of the text is simple''
    \end{itemize}
\end{example}

@@

\begin{example} \ \smallskip
    \begin{itemize}
        \item ``this sentence isn't simple''
        \item ``is this sentence simple?''
        \item ``this sentence was simple''
        \item ``this sentence will be simple''
        \item ``this sentence is simpler than that sentence''
    \end{itemize}
\end{example}

@@

\begin{itemize}
    \item while GF can parse the SL string in a variety of ways and, as a consequence, return multiple TL linearizations, our program only outputs the first candidate translation, preliminarily discarding trees containing terminal nodes that have no linearization in the TL\footnote{a linearization can be missing, for instance, if a concrete syntax is postprocessed by simply getting rid of an ``illegal'' verb causing compilation problems (cf. Section \ref{grgn})} 
    \item to facilitate the experiments, our standalone has been written so that it can translate several newline-separated sentences at once.
\end{itemize}

@@

\begin{figure}[h]
    \centering
    \footnotesize
    \begin{subfigure}{.5\textwidth}
      \centering
      \setlength{\unitlength}{0.2mm}
      \begin{picture}(315.0,110.0)
        \put(0.0,0.0){parse}
        \put(55.0,0.0){this}
        \put(101.0,0.0){sentence}
        \put(183.0,0.0){with}
        \put(229.0,0.0){care}
        \put(0.0,15.0){{\tiny VERB}}
        \put(55.0,15.0){{\tiny DET}}
        \put(101.0,15.0){{\tiny NOUN}}
        \put(183.0,15.0){{\tiny ADP}}
        \put(229.0,15.0){{\tiny NOUN}}
        \put(88.0,30.0){\oval(39.47826086956522,33.333333333333336)[t]}
        \put(68.26086956521739,35.0){\vector(0,-1){5.0}}
        \put(73.0,49.66666666666667){{\tiny det}}
        \put(70.5,30.0){\oval(98.02970297029702,66.66666666666667)[t]}
        \put(119.51485148514851,35.0){\vector(0,-1){5.0}}
        \put(55.5,66.33333333333334){{\tiny obj}}
        \put(216.0,30.0){\oval(39.47826086956522,33.333333333333336)[t]}
        \put(196.26086956521738,35.0){\vector(0,-1){5.0}}
        \put(201.0,49.66666666666667){{\tiny case}}
        \put(134.5,30.0){\oval(227.68995633187774,100.0)[t]}
        \put(248.34497816593887,35.0){\vector(0,-1){5.0}}
        \put(119.5,83.0){{\tiny obl}}
        \put(15.0,110.0){\vector(0,-1){80.0}}
        \put(20.0,100.0){{\tiny root}}
      \end{picture}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \setlength{\unitlength}{0.25mm}
\begin{picture}(324.0,90.0)
  \put(0.0,0.0){analizza}
  \put(82.0,0.0){questa}
  \put(146.0,0.0){frase}
  \put(201.0,0.0){con}
  \put(238.0,0.0){cura}
  \put(0.0,15.0){{\tiny VERB}}
  \put(82.0,15.0){{\tiny DET}}
  \put(146.0,15.0){{\tiny NOUN}}
  \put(201.0,15.0){{\tiny ADP}}
  \put(238.0,15.0){{\tiny NOUN}}
  \put(124.0,30.0){\oval(59.3125,33.333333333333336)[t]}
  \put(94.34375,35.0){\vector(0,-1){5.0}}
  \put(109.0,49.66666666666667){{\tiny det}}
  \put(93.0,30.0){\oval(143.94520547945206,66.66666666666667)[t]}
  \put(164.97260273972603,35.0){\vector(0,-1){5.0}}
  \put(78.0,66.33333333333334){{\tiny obj}}
  \put(229.5,30.0){\oval(28.89189189189189,33.333333333333336)[t]}
  \put(215.05405405405406,35.0){\vector(0,-1){5.0}}
  \put(214.5,49.66666666666667){{\tiny case}}
  \put(212.0,30.0){\oval(88.73913043478261,66.66666666666667)[t]}
  \put(256.3695652173913,35.0){\vector(0,-1){5.0}}
  \put(197.0,66.33333333333334){{\tiny nmod}}
  \put(15.0,90.0){\vector(0,-1){60.0}}
  \put(20.0,80.0){{\tiny root}}
\end{picture}
    \end{subfigure}
    \caption[A case where alignment ``by exclusion'' is beneficial]{A case where alignment ``by exclusion'' is beneficial. In this case, while the two sentences should appear to be structured identically to a human reader, the Italian parse tree contains a (reasonable) attachment error. As a consequence, vanilla CE would not be able to align the prepositional phase ``\textit{with care}'', attached to the main verb ``\textit{parse}'', to its Italian exact counterpart ``\textit{con cura}'', incorrectly linked to the noun ``\textit{frase}'' as an \texttt{nmod}. A second CE pass trying to put unaligned subtrees in relation with each other, however, is able to detect the POS-equivalence between the two phrases, thus producing three new correct alignments: $\langle$\textit{with care, con cura}$\rangle$, $\langle$\textit{``care'', ``cura''}$\rangle$ and $\langle$\textit{``with'', ``con''}$\rangle$.}
    \label{rest}
\end{figure}

@@

On the other hand, the concepts represented in the GF grammar should intuitively, in most cases, be ``as small as possible''. If we reconsider the example in Figure \ref{cake} (sentences \textit{``finding useful correspondences is not exactly a piece of cake''} and \textit{``trovare corrispondenze utili non è proprio scontato''}), for instance, while the multi-word correspondence $\langle$\textit{``a piece of cake'', ``scontato''}$\rangle$ is indeed useful there is usually no point in keeping correspondences such as $\langle$\textit{``useful correspondences'', ``corrispondenze utili''}$\rangle$, even if they are correct, when it it also possible to extract the alignments $\langle$\textit{``useful'', ``utili''}$\rangle$ and $\langle$\textit{``correspondences'', ``corrispondenze''}$\rangle$: one-word and, in general, shorter alignments are more easily reusable in different contexts, especially since there are language-specific RGL rules capable of handling word order properly. 
As a consequence, for the purposes of grammar generation, the CA modules should in this case only return the following set of alignments: \smallskip

\begin{enumerate}
    \item $\langle$\textit{``finding'', ``trovare''}$\rangle$
    \item $\langle$\textit{``useful'', ``utili''}$\rangle$
    \item $\langle$\textit{``correspondences'', ``corrispondenze''}$\rangle$
    \item $\langle$\textit{``is'', ``è''}$\rangle$
    \item $\langle$\textit{``not'', ``non''}$\rangle$
    \item $\langle$\textit{``exactly'', ``proprio''}$\rangle$
    \item $\langle$\textit{``a piece of cake'', ``scontato''}$\rangle$.
\end{enumerate} \smallskip

@@

In some situations, however, larger alignments (such as $\langle$\textit{``useful correspondences'', ``corrispondenze utili''}$\rangle$) are considered to be relevant since they help understand which words can be safely used in conjunction with each other. For instance, while the Italian adjective ``\textit{brutto}'' (most often translated as ``\textit{ugly}'') is often used in conjunction with the word ``\textit{notizia}'' (``\textit{news}'') in expressions such as ``\textit{brutte notizie}'', in English the adjective ``\textit{bad}'' (as in ``\textit{bad news}'') is definitely more appropriate. Discarding a potential $\langle$\textit{``bad news''}, \textit{``brutte notizie''}$\rangle$ alignment is then not always the right choice, since in a context in which there are many alternative translation equivalents for the word ``\textit{brutto}'' it provides useful information about what adjective to use to refer to the noun ``\textit{news}''. \smallskip

@@

Thus, the algorithm is ``safe'' in the sense that it does not get rid of any potentially meaningful alignment. 
In fact, it only gets rid of superficially perfect alignments whose subtrees are also members of smaller alignments. 
This means that when sentences vary widely syntax-wise, as it happens in cases where translation is not literal or in the presence of parse errors, the resulting extracted concepts can still contain a large number of nodes.

@@

In a preliminary experiment conducted in this way, for instance, the Italian translation ``\textit{il docente discute la Didattica}'' of the English sentence ``\textit{the teacher discusses the didactics}'', was given the minimum score despite being completely correct and even idiomatic\footnote{typically, the adjective ``\textit{didattica}'' is not used in its plural form ``\textit{didattiche}'' when nominalized. The capital D is often used, for this particular term, in official university documents.}. 
Looking at the two reference translations used at the time, it appears clear that such a low score is due to the capitalization of the word ``\textit{Didattica}'' and the arbitrary (but equally correct) lexical choices made by the human translators, who rendered the sentence as ``\textit{il professore discute i metodi didattici}'' and ``\textit{l' insegnante discute la didattica}''\footnote{The terms ``\textit{docente}'', ``\textit{professore}'' and ``\textit{insegnante}'' are almost perfect synonyms, even though ``\textit{docente}'' is mostly used in formal documents to refer to university lecturers, ``\textit{professore}'' is more common in secondary school and colloquial speech, and ``\textit{insegnante}'' is the most generic. As such, it can also refer to elementary school teachers and in some cases even sport coaches. The expression ``\textit{metodi didattici}'' is a perfectly sensible replacement for ``\textit{didattica}''.}. 
Coincidentally (but not unlikely for a language like Italian), this is a case in which these lexical differences affect the choice of the determiners used in the various translations (cf. ``\textit{\underline{l'}insegnante}'' vs. ``\textit{\underline{il} professore}'' and ``\textit{\underline{la} didattica}'' vs. ``\textit{\underline{i} metodi didattici}''), producing a BLEU-1 to 4 score of 0. \smallskip

@@

These synthetic figures are useful to give an idea of the general quality of the translations: overall, although with relatively low scores, English-to-Swedish translation works significantly better than English-to-Italian. 
Looking back at the results obtained for CE (cf. Section \ref{eval2}), this is not excessively surprising, since both precision and recall for English-Swedish are consistently higher than they are for English-Italian. 
When it comes to the results obtained for CE on the course plans corpora in particular (cf. Table \ref{tcereal}), however, the difference in precision between the two language pairs is negligible.
This makes the reason for the substantial difference in the BLEU scores obtained for the two different corpora unclear, at least at a first glance.\smallskip

@@

On the other hand, it is easy for shorter sentences to be assigned the minimum BLEU-1 to 4 score even when they only contain a single grammatical or semantic error. 
This is the case, for instance, of the sentence ``\textit{the test is oral}'', whose last word, ``\textit{oral}'' is translated as ``\textit{dura}'' (``\textit{hard}'') instead of ``\textit{orale}'' due to an alignment error. 
Nonetheless, it is worth noticing that this is one of the many cases in which the correct alignment $\langle$``\textit{oral}'',``\textit{orale}''$\rangle$ is also found by the CE module. 
This is hidden by the fact that, for simplicity, only the first translation candidate is taken into account.
Moreover, such valid correspondence has a higher number of occurrences and matches stronger alignment criteria than the wrong one $\langle$``\textit{oral}'',``\textit{duro}''$\rangle$. 
This suggest that, in contexts where higher precision is necessary, there are obvious improvements that can be done either in the choice of the alignment criteria or at a later stage, when relevant alignments are selected. \smallskip

@@

In Italian, grammar errors often involve contractions such as ``\textit{del}'' (``\textit{di}'' + ``\textit{il}'', in English ``\textit{of the}''), some of which are systematically rendered as two separate words.
Another common case is that of wrong (or at least very confusing) adjective collocation, such as in the translation ``\textit{il libro presenta una tecnica con miglioramenti vari utile}'' (``\textit{the textbook presents a useful technique with various improvements}''), where the adjective ``\textit{utile}'' (``\textit{useful}''), referred to ``\textit{technique}'' (``\textit{tecnica}'') is placed far from such noun, making the sentence hard to interpret\footnote{the manually postprocessed translation is ``\textit{il libro presenta una utile tecnica con miglioramenti vari}''.}.
Grammatical errors in Swedish are less common and, apart from long adjectives never being turned to the comparative degree periphrastically (e.g. ``\textit{relevant\underline{are}}'' instead of ``\textit{mer relevant}''), less systematic. 
Only in one case, for instance, gender is incorrect (``\textit{programbibliotek\underline{en}}''). \smallskip

Some errorsregarding the extracted concepts are also interesting to analyze.
In English-Swedish, while several compounds, such as $\langle$\textit{``computer science'', ``datavetenskap''}$\rangle$, are aligned correctly, there are cases in which a single-root noun is translated as a compound (e.g. $\langle$ \textit{``theory'', ``automatateori''} $\rangle$). This is not necessarily an actual alignment error, as it might rather be that the English version of the text was being less specific than its Swedish counterpart, thus producing an alignment that, during evaluation, would have been marked as ``correct but not reusable'' (\texttt{=}). \smallskip

``Reasonable'' alignment errors appear in the DMI corpus sentences too. 
The English phrase ``\textit{attendance of lessons}'', for instance, becomes simply ``\textit{frequenza}'' in the Italian translation. 
Such word is in fact often used alone to replace longer expressions such as ``\textit{frequenza delle lezioni}'', just like in English ``\textit{of lessons}'' can be omitted when what must be attended is evident from the context.
Another interesting example is the alignment $\langle$\textit{``class'', ``classe''}$\rangle$, which causes the sentence \textit{``I will attend the class''} to be (incorrectly) translated as \textit{``io seguirò la classe''} instead of \textit{``io seguirò la lezione''} even though the correspondence is in fact valid in most of the numerous contexts in which \textit{``class''} is not to be intended as a synonym of \textit{``lesson''}.

### 2021-02-09 final_report/include/CP.tex
reviewed MT eval chapter

\begin{table}[H]
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
    \textbf{}              & \textbf{en-sv} & \textbf{it-sv} \\ \hline
    propagated             & 1019 (85.05\%) & 979 (84.64\%) \\ \hline
    tot. errors            & 133 (13.05\%)  & 187 (19.1\%)  \\ \hline
    CP-introduced          & 75 (56.39\%)   & 84 (44.91\%)   \\ \hline
    \end{tabular}
    \caption[Performance of CP Scenario 1 on manually annotated data]{Results of propagating the Italian-English alignments obtained via CE on the first 100 sentences of the PUD corpus to Swedish, using either the English or Italian subtrees as a starting point.}
    \label{tcp1}
    \end{table}

@@

\begin{table}[H]
    \centering
    \scriptsize
    \begin{tabular}{|l|l|l|l|l|l|l|}
    \hline
    \textbf{}                     & \textbf{en-it-sv} & \textbf{it-en-sv} & \textbf{en-sv-it} & \textbf{sv-en-it} & \textbf{it-sv-en} & \textbf{sv-it-en} \\ \hline
    extracted   & 638               & 638               & 687               & 687               & 608               & 608               \\ \hline
    propagated                    & 92 (14.42\%)      & 92 (14.42\%)      & 98 (14.26\%)      & 84 (12.22\%)      & 101 (16.61\%)     & 87 (14.37\%)      \\ \hline
    tot. errors                   & 46 (50\%)         & 21 (22.82\%)      & 42 (42.85\%)      & 24 (28.57\%)      & 21 (20.79\%)      & 28 (32.18\%)      \\ \hline
    CP-introduced                 & 33 (71.73\%)      & 11 (52.38\%)      & 21 (50\%)         & 12 (50\%)         & 12 (57.14\%)      & 21 (75\%)         \\ \hline
    \end{tabular}
    \caption[Performance of CP Scenario 2 on manually annotated non-homogeneous data]{Results of CP for various language pairs using PUD (non-homogeneous) data. For each column, the first two languages are the ones used for extraction and the second two the ones used for propagation.}
    \label{tcp2pud}
    \end{table}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
                           & \textbf{sv-en-it} & \textbf{it-en-sv} \\ \hline
    extracted concepts     & 1950              & 1823              \\ \hline
    propagated             & 205 (10.51\%)     & 200 (10.97\%)     \\ \hline
    tot. errors            & 66 (32.19\%)      & 61 (30.5\%)       \\ \hline
    errors introduced by CP & 33 (50\%)        & 33 (54.09\%)      \\ \hline
    \end{tabular}
    \caption[Performance of CP Scenario 2 on automatically parsed homogeneous data]{Results of CP for using two different bilingual corpora belonging to the same domain of course plans. For each column, the first two languages are the ones used for extraction and the second two the ones used for propagation.}
    \label{tcp2courses}
    \end{table}

>>>>

\begin{table}[H]
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
    \textbf{}              & \textbf{en-sv} & \textbf{it-sv} \\ \hline
    \textbf{propagated}             & 1019 (85.05\%) & 979 (84.64\%) \\ \hline
    \textbf{tot. errors}            & 133 (13.05\%)  & 187 (19.1\%)  \\ \hline
    \textbf{CP-introduced}          & 75 (56.39\%)   & 84 (44.91\%)   \\ \hline
    \end{tabular}
    \caption[Performance of CP Scenario 1 on manually annotated data]{Results of propagating the Italian-English alignments obtained via CE on the first 100 sentences of the PUD corpus to Swedish, using either the English or Italian subtrees as a starting point.}
    \label{tcp1}
    \end{table}

@@

\begin{table}[H]
    \centering
    \scriptsize
    \begin{tabular}{|l|l|l|l|l|l|l|}
    \hline
    \textbf{}                     & \textbf{en-it-sv} & \textbf{it-en-sv} & \textbf{en-sv-it} & \textbf{sv-en-it} & \textbf{it-sv-en} & \textbf{sv-it-en} \\ \hline
    \textbf{extracted}   & 638               & 638               & 687               & 687               & 608               & 608               \\ \hline
    \textbf{propagated}                    & 92 (14.42\%)      & 92 (14.42\%)      & 98 (14.26\%)      & 84 (12.22\%)      & 101 (16.61\%)     & 87 (14.37\%)      \\ \hline
    \textbf{tot. errors}                   & 46 (50\%)         & 21 (22.82\%)      & 42 (42.85\%)      & 24 (28.57\%)      & 21 (20.79\%)      & 28 (32.18\%)      \\ \hline
    \textbf{CP-introduced}                 & 33 (71.73\%)      & 11 (52.38\%)      & 21 (50\%)         & 12 (50\%)         & 12 (57.14\%)      & 21 (75\%)         \\ \hline
    \end{tabular}
    \caption[Performance of CP Scenario 2 on manually annotated non-homogeneous data]{Results of CP for various language pairs using PUD (non-homogeneous) data. For each column, the first two languages are the ones used for extraction and the second two the ones used for propagation.}
    \label{tcp2pud}
    \end{table}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
                           & \textbf{sv-en-it} & \textbf{it-en-sv} \\ \hline
    \textbf{extracted}     & 1950              & 1823              \\ \hline
    \textbf{propagated}             & 205 (10.51\%)     & 200 (10.97\%)     \\ \hline
    \textbf{tot. errors}            & 66 (32.19\%)      & 61 (30.5\%)       \\ \hline
    \textbf{CP-introduced} & 33 (50\%)        & 33 (54.09\%)      \\ \hline
    \end{tabular}
    \caption[Performance of CP Scenario 2 on automatically parsed homogeneous data]{Results of CP for using two different bilingual corpora belonging to the same domain of course plans. For each column, the first two languages are the ones used for extraction and the second two the ones used for propagation.}
    \label{tcp2courses}
    \end{table}
